MCMC Redistricting

### 2020-12-14 paper.Rmd
Results Section

In figure 1 above we see the current district map for the state of colorado. The districts are distinguished by the different colors and the legend on the right. The opaque white lines show the precinct lines (there are over 3000 precincts in the state of Colorado), as mentioned in the intorduction gerrymandering is the process of moving precincts from one district to another to sway political bias in certain districts or in an entire state. Each district shown above satisfies the equal population constraint stating that every district must have approximately the same population. At the moment politicians are doing this process by hand looking at each precinct's voting history and moving them around to benefit their political party. Redistricting happens once every 10 years after every census. The current district layout exhibits democratic bias which is as expected because democrats controlled the state legislature in 2010 which was the last census year.

@@

Now we can look at the bias asscoiated with our simulations. Our simulation consisted of 10,000 different and unique redistricting maps that the algorithm computed using our parameters and equal population constraint. Each point on the graph is a distinct Colorado district map. On the X-axis we have the percentage of precincts swapped for a particular simulation. On the Y-axis we have our calculated political bias. The red dotted line shows 0 bias, above that line is democratic bias and below that line is republican bias. The black line shows the current political bias which is slightly democratic.

There seems to be a connection between political bias and percentage of precincts swapped. When less than 10 percent of precincts are swapped the political bias changes very little and stays democratic but it can sway slightly closer to the ideal 0 bias line. However as more precincts get swapped we start to see a positive trend in the graph where partisan bias grows more and more in favor of democrats. According to our simulations in order to acheive 0 bias approximately 37 percent of precincts need to be swapped from their original district. The political bias can even become slightly republican with this proportion of precincts swapped. However, the over all trend is that the vast meajority of the simulations that were run return a district map of colorado that is even more democratically biased than it currently is. We believe this is due to the fact that Colorado has been trending away from being a swing state and becoming more of a solid "blue" shown by the outcomes of the presedential elections in the past 20 years.

@@

Here is the minimum bias map. This map shows the simulation with the political bias that is closest to 0. This map does satisfy all of the constraints put forth by the state of colorado for redistricting however it may be slightly unrealistic as far as a real world district map. The districts developed a "tree branch pattern" where they seem to branch out accross the entire state of colorado from the center outwards. This is because of a compactness constraint that we were not able to implement due to our limited computing power. The compactness constraint takes in a matrix of pair wise distances between precincts and makes sure that all precincts in a district are not "too" far away from eachother. We tried running our simulations using this constraint but it crashed our computers so we decided to run or simulations without compactness. However we think there is something to be learned from this minimum bias map, it shows just how different districts need to look in Colorado to obtain a 0 bias state.

@@

Moving forward we would like to exlpore the compactness constraint to attempt to get a more realistic redistricted map of Colorado. Also bumping up the number of simulations could prove to be useful, our limit with the computers that we have is 10,000 runs getting that number up to 100,000 would be ideal. However being able to do those two things woiuld require access to more powerful computers. Also running this analysis on another state could be very interesting, Texas for example, a historically republican state, has shown a trend of democratic voting increase. It would be very interesting to look into how gerrymandering has been affecting Texas politics.

>>>>

In figure 1 above we see the current district map for the state of colorado. The districts are distinguished by the different colors and the legend on the right. The opaque white lines show the precinct lines (there are over 3000 precincts in the state of Colorado), as mentioned in the introduction gerrymandering is the process of moving precincts from one district to another to sway political bias in certain districts or in an entire state. Each district shown above satisfies the equal population constraint stating that every district must have approximately the same population. At the moment politicians are doing this process by hand looking at each precinct's voting history and moving them around to benefit their political party. Redistricting happens once every 10 years after every census. The current district layout exhibits democratic bias which is as expected because democrats controlled the state legislature in 2010 which was the last census year.

@@

Now we can look at the bias associated with our simulations. Our simulation consisted of 10,000 different and unique redistricting maps that the algorithm computed using our parameters and equal population constraint. Each point on the graph is a distinct Colorado district map. On the X-axis we have the percentage of precincts swapped for a particular simulation. On the Y-axis we have our calculated political bias. The red dotted line shows 0 bias, above that line is democratic bias and below that line is republican bias. The black line shows the current political bias which is slightly democratic.

There seems to be a connection between political bias and percentage of precincts swapped. When less than 10 percent of precincts are swapped the political bias changes very little and stays democratic but it can sway slightly closer to the ideal 0 bias line. However as more precincts get swapped we start to see a positive trend in the graph where partisan bias grows more and more in favor of democrats. According to our simulations in order to achieve 0 bias approximately 37 percent of precincts need to be swapped from their original district. The political bias can even become slightly republican with this proportion of precincts swapped. However, the overall trend is that the vast majority of the simulations that were run return a district map of colorado that is even more democratically biased than it currently is. We believe this is due to the fact that Colorado has been trending away from being a swing state and becoming more of a solid "blue" shown by the outcomes of the presidential elections in the past 20 years.

@@

Here is the minimum bias map. This map shows the simulation with the political bias that is closest to 0. This map does satisfy all of the constraints put forth by the state of colorado for redistricting however it may be slightly unrealistic as far as a real world district map. The districts developed a "tree branch pattern" where they seem to branch out across the entire state of Colorado from the center outwards. This is because of a compactness constraint that we were not able to implement due to our limited computing power. The compactness constraint takes in a matrix of pairwise distances between precincts and makes sure that all precincts in a district are not "too" far away from each other. We tried running our simulations using this constraint but it crashed our computers so we decided to run our simulations without compactness. However we think there is something to be learned from this minimum bias map, it shows just how different districts need to look in Colorado to obtain a 0 bias state.

@@

Moving forward we would like to explore the compactness constraint to attempt to get a more realistic redistricted map of Colorado. Also bumping up the number of simulations could prove to be useful, our limit with the computers that we have is 10,000 runs getting that number up to 100,000 would be ideal. However being able to do those two things would require access to more powerful computers. Also, running this analysis on another state could be very interesting, Texas for example, a historically republican state, has shown a trend of democratic voting increase. It would be very interesting to look into how gerrymandering has been affecting Texas politics.
